---
title: "R4CSR"
format: revealjs
editor: visual
---

# Book club

This is a slide deck summary of [R for Clinical Study Reporting and Submission](https://r4csr.org/)

There are two main parts

1.  TLFs: a simulation of individual contributions

2.  Clinical trial projects: a simulation of project leadership

[Join us!](https://datascience.arizona.edu/events/r-clinical-study-reporting)

## Get R help at UArizona

[Data & Visualization Drop-in](https://datascience.arizona.edu/events/data-viz-drop-0) hours @ Main library

## Preface

-   R folder structure recommended

-   [CDISC pilot data](https://github.com/cdisc-org/sdtm-adam-pilot-project/tree/master/updated-pilot-submission-package/900172/m5/datasets/cdiscpilot01/analysis/adam/datasets) will be used

-   R packages needed

```{r}
#| eval: false
#| echo: true

install.packages(c(
                    "dplyr"   # Manipulate data
                  , "r2rtf" # Reporting in RTF format
                  , "tidyr" # Manipulate data
                       ))
```

## Living list of acronyms {.scrollable}

| Type                   | Acronym | Definition/Explanation                                                                                                          |
|----------------|----------------|----------------------------------------|
| clinical               | CSR     | Clinical study report                                                                                                           |
| clinical               | SDTM    | standard data tabulation model                                                                                                  |
| clinical               | ADaM    | Analysis dataset model                                                                                                          |
| clinical               | TLFs    | Tables, listings, and figures                                                                                                   |
| clinical               | A&R     | Analysis and reporting                                                                                                          |
| clinical/computational | eCTD    | Electronic common technical document                                                                                            |
| clinical               | CDISC   | Clinical Data Interchange Standards Consortium                                                                                  |
| clinical               | ICH     | International conference on harmonization                                                                                       |
| clinical               | ICH E3  | ICH guidelines on structure and content of clinical study reports                                                               |
| computational          | RTF     | Rich text format                                                                                                                |
| clinical               | adae    | Analysis dataset for adverse events                                                                                             |
| clinical/computational | ADSL    | Subject-level analysis dataset                                                                                                  |
| clinical/computational | ITT     | intention to treat (i.e. none of the patients are excluded and the patients are analyzed according to the randomization scheme) |

# TLFs {style="text-align: center;"}

# Chapter 1

An overview of clinical study reports and the {r2rtf} package.

## Overview

-   A CSR contains all methods and results of a study in \~16 sections

    -   TLFs are found in 10-12, 14, 16, +

-   often word document format

    -   in this book we explore RTF

-   The {r2rtf} package allows for the creation and export of publication quality tables and figures in rich text format

    -   Use the included `r2rtf_adae` dataset to explore functions in {r2rtf}

## r2rtf {#r2rtf}

After formatting your data as desired by {r2rtf}, a general workflow is as follows:

```{r}
#| echo: true
#| eval: false

head(tbl) %>%
  rtf_body() %>% # Step 1 Add table attributes
  rtf_encode() %>% # Step 2 Convert attributes to RTF encode
  write_rtf("tlf/intro-ae1.rtf") # Step 3 Write to a .rtf file
```

Each table attribute is added individually, then the attributes are converted to RTF, and finally, you can export an object in RTF.

## Style with r2rtf

Use {r2rtf} to set specific design elements for your ouput:

```{r}
#| echo: true
#| eval: false

head(tbl) %>%
  rtf_colheader(
    colheader = " | Treatment",
    col_rel_width = c(3, 6)
  ) %>%
  rtf_colheader(
    colheader = "Adverse Events | Placebo | Xanomeline High Dose | Xanomeline Low Dose",
    border_top = c("", "single", "single", "single"),
    col_rel_width = c(3, 2, 2, 2)
  ) %>%
  rtf_body(col_rel_width = c(3, 2, 2, 2)) %>%
  rtf_encode() %>%
  write_rtf("tlf/intro-ae7.rtf")
```

A single {r2rtf} command may include columns, borders, headers, width designations and many other elements.

# Chapter 2

Section 10.1, Disposition of Participants

> The disposition of participants table reports the numbers of participants who were randomized, and who entered and completed each phase of the study, and the reasons for all post-randomization discontinuations, grouped by treatment and by major reason (lost to follow-up, adverse event, poor compliance, etc.) are reported.

## Disposition

Step 1: Count participants in the analysis population

Step 2: Calculate the number and percentage of participants who discontinued the study by treatment arm

Step 3: Calculate the numbers and percentages of participants who discontinued the study for different reasons, by treatment arm

Step 4: Calculate the number and percentage of participants who completed the study, by treatment arm

Step 5: Bind `n_rand`, `n_disc`, `n_reason`, and `n_complete` by row.

Step 6+: Write the final table to RTF

## Disposition of participants RTF

::: panel-tabset
### Code

```{r}
#| eval: false
#| echo: true

tbl_disp %>%
  # Table title
  rtf_title("Disposition of Participants") %>%
  # First row of column header
  rtf_colheader(" | Placebo | Xanomeline Low Dose| Xanomeline High Dose",
    col_rel_width = c(3, rep(2, 3))
  ) %>%
  # Second row of column header
  rtf_colheader(" | n | (%) | n | (%) | n | (%)",
    col_rel_width = c(3, rep(c(0.7, 1.3), 3)),
    border_top = c("", rep("single", 6)),
    border_left = c("single", rep(c("single", ""), 3))
  ) %>%
  # Table body
  rtf_body(
    col_rel_width = c(3, rep(c(0.7, 1.3), 3)),
    text_justification = c("l", rep("c", 6)),
    border_left = c("single", rep(c("single", ""), 3))
  ) %>%
  # Encoding RTF syntax
  rtf_encode() %>%
  # Save to a file
  write_rtf("tlf/tbl_disp.rtf")
```

### Comments

With our bound data we follow our [workflow for {r2rtf}](#r2rtf): add attributes, convert to RTF, write out

-   `|` separates every item, thus line 5 denotes 4 columns and line 9, 7 cols
-   line 6 could also read\
    `col_rel_width = c(3, 2, 2, 2)`
-   and line 10 could also read\
    `col_rel_width = c(3, 0.7, 1.3, 0.7, 1.3, 0.7, 1.3)`
:::

# Chapter 3

Section 11.1, Data Sets Analyzed

> The summary of analysis sets table reports on participants included in each efficacy analysis

## Writing functions

> \"You should consider writing a function whenever you\'ve copied and pasted a block of code more than twice\"
>
> \- [R for Data Science](https://r4ds.had.co.nz/functions.html#when-should-you-write-a-function)

\

```{r}
#| eval: false
#| echo: true


fmt_num <- function(x, digits, width = digits + 4) {
  formatC(x,
    digits = digits,
    format = "f",
    width = width
  )
}
```

## Summary of analysis sets

With helper functions `count_by` and `fmt_num` we can more easily prepare the dataset for a summary of analysis sets table with the following steps:\
\

Step 1: Bind the counts/percentages of the ITT population, the efficacy population, and the safety population by row using the `count_by()` function. \
Step 2+: Format the output from Step 2 using r2rtf.

# Clinical Trial Project (June 15, 2023) {style="text-align: center;"}
